---
title: "imager FAQ"
output:
  html_document:
    fig_width: 6.5
    fig_height: 6
    toc: true
    number_sections: true
---

[Simon Barthelmé](http://sites.google.com/site/simonbarthelme) (GIPSA-lab, CNRS)

# I can't install the package

If you're on OS X, you're probably missing XQuartz: https://www.xquartz.org/. 
Otherwise, you're probably attempting to compile from source. If you're installing from github, try installing from CRAN (install.packages('imager')).
If CRAN can't find binaries for your platform:
- If you're on OS X, make sure you have installed Xcode. 
- If you're on Linux, make sure you have all the build essentials (installing libfftw3-dev and g++ should pull all the dependencies you need). 

# I can't load a file/video 

imager by itself can only load images in png, jpeg and bmp format. Install [ImageMagick](http://www.imagemagick.org/script/binary-releases.php) for more. 
If you want to load videos, install [ffmpeg](http://ffmpeg.org/download.html). 

# I tried loading a video and my computer froze

imager doesn't do out-of-memory processing, meaning the *whole video* gets loaded. Videos take up huge amounts of memory. If you really want to muck around with large videos, try the experimental package [imagerstreams](https://github.com/dahtah/imagerstreams), which supports out-of-memory processing. 

# How do I access file metadata? 

Try iminfo (needs imagemagick).

# How do I overlay/stack images? 

For overlaying, see imdraw:

```{r}
load.example("parrots") %>% imdraw(boats/255,opacity=.5) %>% plot
```

For stacking, use imappend:

```{r}
a <- load.example("parrots")
b <- boats/255
list(a,b) %>% imappend("x") %>% plot
```

For more flexibility, you can also pad the image and use imdraw. 
