---
title: "imager: an R package for image processing"
output:
  html_document:
    toc: true
    number_sections: true
---    

[Simon Barthelmé](http://sites.google.com/site/simonbarthelme) (GIPSA-lab, CNRS)


```{r init,echo=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE, cache=FALSE, 
               comment=NA, verbose=TRUE,dev='jpeg')
```

This documentation covers imager version 0.40. Some functions may be unavailable in older versions. Follow imager development on [github](https://github.com/dahtah/imager).

Beginners: have a look at the [tutorial](gettingstarted.html) first. 

# Quick start

Here's an example of imager in action:

```{r quickstart,message=FALSE,cache=TRUE}
library(imager)


file <- system.file('extdata/parrots.png',package='imager')
#system.file gives the full path for a file that ships with a R package
#if you already have the full path to the file you want to load just run:
#im <- load.image("/somedirectory/myfile.png")
im <- load.image(file)

plot(im) #Parrots!
im.blurry <- isoblur(im,10) #Blurry parrots!
plot(im.blurry)
im.xedges <- deriche(im,2,order=2,axis="x") #Edge detector along x-axis
plot(im.xedges)
im.yedges <- deriche(im,2,order=2,axis="y") #Edge detector along y-axis
plot(im.yedges)
#Chain operations using the pipe operator (from magrittr)
deriche(im,2,order=2,axis="x") %>% deriche(2,order=2,axis="y") %>% plot
```

```{r quickstart2,fig.width=12,cache=TRUE}
#Another example of chaining: image gradient along x and y axes
layout(matrix(1:2,1,2));
grayscale(im) %>% imgradient %>% plot

#If ffmpeg is present, we can load videos as well:
tennis <- load.video(system.file('extdata/tennis_sif.mpeg',package='imager'))

plot(tennis,frame=1)
plot(tennis,frame=5)
```

In the next example, we convert the video to grayscale, run a motion detector, and combine both videos to display them side-by-side:

```{r motion_example,cache=TRUE}
tennis.g <- grayscale(tennis)
motion <- deriche(tennis.g,1,order=1,axis="z")^2 #Differentiate along z axis and square
combined <- list(motion/max(motion),tennis.g/max(tennis.g)) %>% imappend("x") #Paste the two videos together
```

In an interactive session you can run play(combined) to view the results. 

```{r animate,echo=FALSE,fig.show="animate",interval=.03,fig.width=12,cache=TRUE}
imsplit(combined,"z") %>% l_ply(plot)
```

# Loading and saving images

Use load.image and save.image. imager supports PNG, JPEG and BMP natively. If you need to access images in other formats you'll need to install [ImageMagick](http://www.imagemagick.org/script/binary-releases.php).

```{r load_save}
file <- system.file('extdata/parrots.png',package='imager')
parrots <- load.image(file)
#The file format is defined by the extension. Here we save as JPEG
imager::save.image(parrots,"/tmp/parrots.jpeg")
#We call imager::save.image to avoid ambiguity, as base R already has a save.image function 
```

You can load files from your hard drive or from a URL. Loading from URLs is useful when scraping web pages, for example. The following piece of code searches for pictures of parrots (using the rvest package), then loads the first four pictures it finds:

```{r loading_parrots}
library(rvest)
#Run a search query (returning html content)
search <- read_html("https://www.google.com/search?site=&tbm=isch&q=parrot")

#Grab all <img> tags, get their "src" attribute, a URL to an image
urls <- search %>% html_nodes("img") %>% html_attr("src") #Get urls of parrot pictures

#Load the first four, return as image list, display
map_il(urls[1:4],load.image) %>% plot
```

# Loading and saving videos 

If you need to load and save videos please install [ffmpeg](http://ffmpeg.org/download.html) for videos. 

To load videos, use load.video:

```{r}
fname <- system.file('extdata/tennis_sif.mpeg',package='imager')
tennis <- load.video(fname)
```

Use *skip.to* to set the initial frame, and *frames* to set the number of frames to grab: 

```{r}
#Skip to time = 2s, and grab the next 3 frames
vid <- load.video(fname,skip.to=2,frames=4)
#Display individual frames (see imsplit below)
imsplit(vid,"z") %>% plot
```

Use *fps* to set the frame acquisition rate:

```{r}
vid <- load.video(fname,fps=1) #1 frame per sec
#The result is 6 frames, one per sec
imsplit(vid,"z") %>% plot
```

To save videos, use save.video:

```{r}

f <- tempfile(fileext=".avi")
save.video(vid,f)
load.video(f) %>% imsplit("z") %>% plot
unlink(f)

```

One problem with videos is that they very quickly won't fit in memory anymore, so you might have to deal with them piecewise. make.video lets you make a video from a directory containing images representing individual frames:

```{r}

dd <- tempdir()
iml <- map_il(seq(0,20,l=60),~ isoblur(boats,.))
#Generate file names
fnames <- sprintf("%s/image-%i.png",dd,1:length(iml))
#Save each frame in a different file
invisible(purrr::map2(iml,fnames,function(im,f) imager::save.image(im,f)))
f <- tempfile(fileext=".avi")

make.video(dd,f)

```


```{r echo=FALSE}
iml <- map_il(seq(0,20,l=60),~ isoblur(boats,.))
save.video(iml,"increasing_blur.webm")
```

<video width="500"  controls loop>
<source src="imager_files/figure-html/increasing_blur.webm" />
<p>
video of increasing blur levels
</p>



See also the [animation](https://cran.r-project.org/web/packages/animation/index.html) package for more along these lines. 

# Displaying images and videos 

To get a standard R plot use the plot function:

```{r plotting}
plot(parrots)
plot(tennis,frame=1) 
```

In addition imager provides display() (for images) and play() (for videos), which are much faster C++ functions for quickly viewing your results. 
If you'd like to use an external viewer, you can save the image to a temporary file:

```{r tmp_file,cache=TRUE}
tmp <- tempfile(fileext=".png") #Open temp. file
imager::save.image(boats,tmp) #Save image to temp. file
#Call "eog [temp file]" to open in external viewer (Linux only)
system2("eog",tmp)
#On a mac: try system2("open",tmp)
unlink(tmp) #Delete temp. file
```


# How images are represented

Images are represented as 4D numeric arrays, which is consistent with CImg's storage standard (it is unfortunately inconsistent with other R libraries, like spatstat, but converting between representations is easy). 
The four dimensions are labelled x,y,z,c. The first two are the usual spatial dimensions, the third one will usually correspond to depth or time, and the fourth one is colour. Remember the order, it will be used consistently in imager.
If you only have grayscale images then the two extra dimensions are obviously pointless, but they won't bother you much. Your objects will still be officially 4 dimensional, with two trailing flat dimensions.
Pixels are stored in the following manner: we scan the image beginning at the upper-left corner, along the x axis. Once we hit the end of the scanline, we move to the next line. Once we hit the end of the screen, we move to the next frame (increasing z) and repeat the process. If we have several colour channels, then once we're done with the first colour channel we move to the next one.
All in all the different dimensions are represented in the x,y,z,c order. In R the object is represented as a 4D array. Here's an example with a grayscale image:
```{r dim_gray,cache=TRUE}
parrots <- load.example('parrots')
gray.parrots <- grayscale(parrots)
dim(gray.parrots)
```

and a colour image:
```{r dim_colour,cache=TRUE}
dim(parrots)
```

and finally a video, also in colour:

```{r dim_video,cache=TRUE}
dim(tennis)
```

# Coordinates

CImg uses standard image coordinates: the origin is at the top left corner, with the x axis pointing right and the y axis pointing *down*. imager uses the same coordinate system, except the origin is now (1,1) and not (0,0) (the reason being that R indices start at 1 and not at 0).
The number of pixels along the x axis is called the width, along the y axis it's height, along the z axis it's depth and finally the number of colour channels is called "spectrum".

```{r dimensions,cache=TRUE}
width(parrots)
height(parrots)
depth(parrots)
spectrum(parrots)
```


# The cimg class

Imager uses the "cimg" class for its images. "cimg" is just a regular 4d array with an S3 class tacked on so we can have custom plot, print, etc. 
To promote an array to a "cimg" object, use as.cimg:

```{r as.cimg.array,cache=TRUE}
noise <- array(runif(5*5*5*3),c(5,5,5,3)) #5x5 pixels, 5 frames, 3 colours. All noise
noise <- as.cimg(noise)
```

You can treat the object as you would any other array:

```{r math_array,cache=TRUE}
#Arithmetic
sin(noise) + 3*noise 

#Subsetting
noise[,,,1] #First colour channel
dim(noise[1:4,,,] )

```


which makes life easier if you want to use ggplot2 for plotting (see [tutorial](getting_started.html)).

You can also convert matrices to cimg objects:

```{r}
matrix(1,10,10) %>% as.cimg
```

and vectors:

```{r}
1:9 %>% as.cimg
```

which tries to guess what sort of image dimensions you want (see [tutorial](getting_started.html)).

The reverse is possible as well: if you have a data.frame with columns x,y,z,cc,value, you can turn it into a cimg object:

```{r as.cimg.data.frame}
df <- expand.grid(x=1:10,y=1:10,z=1,cc=1)
mutate(df,value=cos(sin(x+y)^2)) %>% as.cimg
```

By default as.cimg.data.frame will try to guess image size from the input. You can also be specific by setting the "dims" argument explicitly:

```{r as.cimg.data.frame2}
mutate(df,value=cos(sin(x+y)^2)) %>% as.cimg(dims=c(10,10,1,1))
```

The reverse is possible as well: you can convert a cimg object to a data.frame

```{r as.df,cache=TRUE}
head(as.data.frame(parrots))
```

or to an array, vector or matrix

```{r}
head(as.array(parrots))
head(as.vector(parrots))
grayscale(parrots) %>% as.matrix %>% dim
```

# Image lists 

Many functions in imager produce lists of image as output (see below). These are given the "imlist" class, e.g. imgradient returns: 

```{r}
imgradient(parrots,"xy") %>% class
```

The imlist class comes with a few convenience functions, for example:

```{r}
imgradient(parrots,"xy") %>% plot
```

and: 

```{r}
imgradient(parrots,"xy") %>% as.data.frame %>% head
```

where the "im" column indexes the image in the list. 

To make an image list directly, use "imlist" or as.imlist

```{r}
imlist(a=parrots,b=3*parrots)
list(a=parrots,b=3*parrots) %>% as.imlist
```

# Pixel sets (pixsets)

Another important datatype in imager is the pixel set (AKA pixset, introduced in imager v0.40). A pixset is a set of pixels, represented as a binary image, and that's what you get when you test properties on images, e.g.: 

```{r}
grayscale(parrots) > .8
```

Internally a pixset is just a array of logicals, so it's no different from what you'd get from running a test on a vector, e.g.: 

```{r}
1:4  >= 2
```

Compared to logical arrays, however, pixsets come with many convenience functions, for plotting, splitting, morphological operations, etc. 

As an example, the following bit of code generates random shapes by filtering & thresholding a random image, splitting it into connected components, and removing the smaller ones:

```{r}
set.seed(1)
imnoise(100,100) %>% isoblur(3) %>%
    threshold(0) %>% split_connected %>%
    purrr::keep(~ sum(.) > 200) %>% plot

```


Pixsets are covered elsewhere, in the vignette ``` vignette('pixsets') ```, and in the [morphology tutorial](morphology.Rmd). 


# Splitting and concatenating images

One often needs to perform separate computations on each channel of an image, or on each frame, each line, etc. This can be achieved using a loop or more conveniently using imsplit:

```{r imsplit}
imsplit(parrots,"c") #A list with three elements corresponding to the three channels
imsplit(parrots,"c") %>% laply(mean) #Mean pixel value in each channel
imsplit(parrots,"x") %>% laply(mean) %>% head #Mean pixel value in each line (across all channels)

##For a more modern version, using purrr
imsplit(parrots,"c") %>% purrr::map_dbl(mean) #Mean pixel value in each channel

```

imsplit contains an additional argument, called "nb". When "nb" is positive imsplit cuts the image into "nb" chunks, along the axis "axis":

```{r}
imsplit(parrots,"x",4) %>% plot
imsplit(parrots,"y",3) %>% plot
```

When "nb" is negative, "nb" pixels defines the chunk size:

```{r}
imsplit(parrots,"x",-250) %>% plot
```

That last panel looks crummy because it's a long-and-thin image stretched into a square. 

The inverse operation to imsplit is called imappend: it takes a list of images and concatenates them along the dimension of your choice.

```{r imappend}
#Sample functions and turn them into separate R,G,B channels
R <- as.cimg(function(x,y) sin(cos(3*x*y)),100,100)
G <- as.cimg(function(x,y) sin(cos(3*x*y + pi/2)),100,100)
B <- as.cimg(function(x,y) exp(-.03*x),100,100)
trippy <- imappend(list(R,G,B),"c") #Bind the three channels into one image
plot(trippy)
```

# Split, apply, combine

Often what one wants to do is to split the image along a certain axis (e.g. colour), apply a transformation separately and recombine the result. iiply does that:
```{r iiply}
iiply(parrots,"c",function(v) v/max(v))  %>% plot

#Same thing but longer:
#imsplit(parrots,"c") %>% llply(function(v) v/max(v) %>% imappend("c") %>% plot
```

The code above separates colour channels, applies a normalisation and recombines the result into an image.
Following the same convention used by plyr, imager also defines ilply (which splits, applies and returns a list), idply (which splits, applies and returns a data.frame) and liply (which applies, combines and returns an image). 

Another way to combine after a split is to take the mean result, or a product, etc (the same way layers are combined in an image editing program). You can do this using shortcuts defined by imager:

```{r add_average_etc,cache=TRUE}
imsplit(parrots,"c") %>% add %>% plot(main="Adding colour channels")

#Use different levels of blur on the same image
blur.layers <- map_il(seq(1,15,l=5),~ isoblur(parrots,.))

blur.layers %>% parmin %>% plot(main="Min across blur levels")
blur.layers %>% parmax %>% plot(main="Max across blur levels")
blur.layers %>% average %>% plot(main="Average across blur levels")
blur.layers %>% parmed %>% plot(main="Median across blur levels")


```

# Sub-images, pixel neighbourhoods, etc.

If you need to select a part of an image, you can use imsub, which is best explained by example:

```{r imsub} 
imsub(parrots,x < 30) #Only the first 30 rows
imsub(parrots,y < 30) #Only the first 30 rows
imsub(parrots,x < 30,y < 30) #First 30 columns and rows
imsub(parrots, sqrt(x) > 8) #Can use arbitrary expressions
imsub(parrots,x > height/2,y > width/2)  #height and width are defined based on the image
imsub(parrots,cc==1) #Colour axis is "cc" not "c" here because "c" is an important R function
```
Pixel neighbourhoods (for example, all nearest neighbours of pixel (x,y,z)) can be selected using stencils. See ?get.stencil and the [vignette](image_statistics.html) on natural image statistics for more.

If you need to access a specific colour channel, use any of the following:

```{r}
R(parrots) 
G(parrots)
B(parrots)
#R(parrots) is equivalent to channel(parrots,1)

#Set all channels to 0 except red
parrots.cp <- load.example("parrots")
G(parrots.cp) <- 0
B(parrots.cp) <- 0
plot(parrots.cp)
```

If you need to access a specific frame, use frame:

```{r}
frame(tennis,1)
#Blur frame 1
frame(tennis,1) <- isoblur(frame(tennis,1),10)
```


If you need pixel values along rows and columns use:

```{r}
imrow(R(parrots),10) %>% plot(main="Red channel along 10th row",type="l")
imcol(B(parrots),10) %>% plot(main="Blue channel along 10th line",type="l")
```

Individual pixels can be accessed using *at* and *color.at*:

```{r}
at(parrots,x=20,y=20,cc=1:3)
color.at(parrots,x=20,y=20)
```

Finally all of this is available under the familiar form of the array subset operator, which tries to save you some typing by filling in flat dimensions: 

```{r}
im <- imfill(4,4)
dim(im) #4 dimensional, but the last two dimensions are singletons
im[,1,,] <- 1:4 #Assignment the standard way
im[,1] <- 1:4 #Shortcut
```


# Denoising

Denoising can be performed using basic filters that average over space:

```{r denoise_blur}
birds <- load.image(system.file('extdata/Leonardo_Birds.jpg',package='imager'))
birds.noisy <- (birds + .5*rnorm(prod(dim(birds)))) 
layout(t(1:2))
plot(birds.noisy,main="Original")
isoblur(birds.noisy,5) %>% plot(main="Blurred")
```

Blurring removes some of the noise but also blurs away the contours. CImg provides an anisotropic blur that does not have that problem:

```{r denoise_aniso}
layout(t(1:2))
plot(birds.noisy,main="Original")
blur_anisotropic(birds.noisy,ampl=1e3,sharp=.3) %>% plot(main="Blurred (anisotropic)")
```



# Colour spaces

To convert from RGB to HSL/HSV/HSI/YUV/YCbCR, run RGBto[...], as in the following example:

```{r hsl,im,fig.width=18}
parrots.hsl <- RGBtoHSL(parrots)
chan <- channels(parrots.hsl) #Extract the channels as a list of images
names(chan) <- c("H","S","L")
#Plot
layout(matrix(1:3,1,3))
l_ply(names(chan),function(nm) plot(chan[[nm]],main=nm))
```

The reverse operation is done by running [...]toRGB. Note that all display functions assume that your image is in RGB. 

```{r trippy_is_back}
YUVtoRGB(trippy) %>% plot
```

If you have a colour image, you convert it to grayscale using the grayscale function. If you have a grayscale image, add colour channels using add.colour:

```{r add.colour}
grayscale(parrots) %>% spectrum
#Image has only one channel (luminance)

grayscale(parrots) %>% add.colour %>% spectrum
#Image is still in gray tones but has R,G,B channels 
```


# Resizing, rotation, etc.

Functions for resizing and rotation should be fairly intuitive:

```{r resize_rotate}
thmb <- resize(parrots,round(width(parrots)/10),round(height(parrots)/10))
plot(thmb,main="Thumbnail") #Pixellated parrots

#Same as above: negative arguments are interpreted as percentages
thmb <- resize(parrots,-10,-10)

imrotate(parrots,30) %>% plot(main="Rotating")
imshift(parrots,40,20) %>% plot(main="Shifting")
imshift(parrots,100,100,boundary=1) %>% plot(main="Shifting (Neumann boundaries)")
imshift(parrots,100,100,boundary=2) %>% plot(main="Shifting (circular)")
```

You can pad an image using "pad":

```{r pad}
pad(parrots,axes="y",140) %>% plot
pad(parrots,axes="y",140,pos=-1) %>% plot
```

autocrop will remove any extra padding:
```{r autocrop}
#The argument to autocrop is the colour of the background it needs to remove
pad(parrots,axes="y",140,pos=-1) %>% autocrop(c(0,0,0)) %>% plot

```


# Warping

Warping maps the pixels of the input image to a different location in the output. Scaling is a special case of warping, so is shifting. Warping relies on a map:
$M(x,y) = (x',y')$

that describes where to send pixel (x,y). Shifting the image corresponds to adding a constant to the coordinates:
$M(x,y) = (x+\delta_x,y+\delta_y)$

In imager:

```{r warp_shift}
map.shift <- function(x,y) list(x=x+10,y=y+30)
imwarp(parrots,map=map.shift) %>% plot
```

The map function should take (x,y) as arguments and output a named list with values (x,y).

The warping algorithm has two modes, "forward" and "backward". In forward mode you go through all $(x,y)$ pixels in the source, and paint the corresponding location $M(x,y)$ in the target image. This may result in unpainted pixels, as in the following example:

```{r warp_scale_forward,cache=TRUE}
map.scale <- function(x,y) list(x=1.5*x,y=1.5*y)
imwarp(parrots,map=map.scale) %>% plot(main="Forward mode")
```

In backward mode you go through all pixels $(x',y')$ in the target image, and look up their ancestor $M^{-1}(x',y')$ in the source image. Backward mode has no missing pixel problems, but now you need to define the inverse map and set the "direction" argument to "backward".

```{r warp_scale_backward,cache=TRUE}
map.scale.bw <- function(x,y) list(x=x/1.5,y=y/1.5)
imwarp(parrots,map=map.scale.bw,direction="backward") %>% plot(main="Backward mode")
```

Of course shifting and scaling things is boring and the whole point of warping is to do things like that:

```{r warped_warp,cache=TRUE}
map <- function(x,y) list(x=exp(y/600)*x,y=y*exp(-sin(x/40)))
imwarp(parrots,map=map,direction="forward") %>% plot()

```

See ?imwarp for more. Note that 3D warping is possible as well. 


# Lagged operators

To compute the difference between successive images in a video, you can use the shift operator:

```{r lag,cache=TRUE}
#Compute difference between two successive frames (at lag 1)
(imshift(tennis,delta_z=1)-tennis) %>% plot(frame=2,main="Difference betw. frames 2 and 1")

#Compute difference between frames (at lag 3)
(imshift(tennis,delta_z=3)-tennis) %>% plot(frame=4,main="Difference between frames 3 and 1")

#note that shift uses interpolation. that makes it relatively slow, but one advantage is that it allows non-integer lags:
#shift(tennis,delta_z=3.5)-tennis
#is valid

```

# Filtering

imager has the usual correlate and convolve operations:

```{r }
flt <- as.cimg(matrix(1,4,4)) #4x4 box filter
grayscale(parrots) %>% correlate(flt) %>% plot(main="Filtering with box filter")

#Here the filter is symmetrical so convolution and correlation should be the same. Check:
a <- grayscale(parrots) %>% correlate(flt)
b <- grayscale(parrots) %>% imager::convolve(flt)
all.equal(a,b)
```

CImg includes  fast implementations of Gaussian (and derivative-of-Gaussian) filters. They are available via the "deriche" and "vanvliet" functions.


```{r fig.width=12}
im <- grayscale(parrots)
layout(t(1:2))
deriche(im,sigma=4,order=2,axis="y") %>% plot(main="2nd deriv of Gaussian along y")
vanvliet(im,sigma=4,order=2,axis="y") %>% plot(main="2nd deriv of Gaussian along y")
```

The Vanvliet-Young filter is typically a better approximation. We can establish this by looking at the impulse response (which should be Gaussian). Here's the one-dimensional case:

```{r}
n <- 1e3
xv <- seq(0,1,l=n) #1D Grid
imp <- imdirac(c(n,1,1,1),n/2,1) #1D signal: Impulse at x = n/2
sig <- 80
#impulse response of the Deriche filter
imp.dr <- deriche(imp,sigma=sig) %>% as.vector
#impulse response of the Vanvliet-Young filter
imp.vv <- vanvliet(imp,sigma=sig) %>% as.vector 
imp.true <- dnorm(xv,sd=sig/n,m=.5) #True impulse response
plot(xv,imp.true/max(imp.true),type="l",lwd=2,xlab="x",ylab="Impulse response")
lines(xv,imp.vv/max(imp.vv),col="blue",lwd=2)
lines(xv,imp.dr/max(imp.dr),col="red",lwd=2)
```

The ideal filter is in black, the Vanvliet-Young filter in blue, the Deriche filter in red. Vanvliet-Young is clearly more accurate, but slightly slower:

```{r}
im <- imfill(3e3,3e3)
system.time(deriche(im,3))
system.time(vanvliet(im,3))
```

In both cases computation time is independent of filter bandwidth, which is a very nice feature (the filters are IIR). 


# FFTs and the periodic/smooth decomposition 

FFTs can be computed via the FFT function:

```{r FFT,cache=TRUE}
im <- as.cimg(function(x,y) sin(x/5)+cos(x/4)*sin(y/2),128,128)
ff <- FFT(im)
plot(ff$real,main="Real part of the transform")
plot(ff$imag,main="Imaginary part of the transform")
sqrt(ff$real^2+ff$imag^2) %>% plot(main="Power spectrum")
```

If you want to use CImg's FFT on images of arbitrary size you should enable FFTW3 support. Install FFTW3 on your system, then run install_github("dahtah/imager",ref="fftw"). As a workaround you can also use R's native fft:

```{r FFT_nativeR,cache=TRUE}
rff <- as.matrix(im) %>% fft
Re(rff) %>% as.cimg %>% plot(main="Real part of the transform (R's native code)")
```

Important: both FFT and fft will attempt to perform a multi-dimensional FFT of the input, with dimensionality defined by the dimensionality of the array. If you want to compute a 2D FFT for every frame of a video, use a split (imsplit or ilply).

The FFT works best for periodic signals. One way of making signals periodic is via zero-padding (use the pad function), another is to use the periodic-smooth decomposition of Moisan (2011):

```{r periodicsmooth,fig.width=12,cache=TRUE}
layout(t(1:2))
periodic.part(parrots) %>% plot(main="Periodic part of parrots")
(parrots- periodic.part(parrots)) %>% plot(main="Smooth residual of parrots")
```

See ?periodic.smooth for details.

# Morphology

See the vignette on  [morphology](morphology.html). 
