---
title: "Morphological operations in imager"
output:
  html_document:
    fig_width: 6.2
    fig_height: 8
    toc: true
    number_sections: true
---

```{r init,echo=FALSE}
library(imager)
```
[Simon Barthelmé](http://sites.google.com/site/simonbarthelme) (GIPSA-lab, CNRS)


Our first example will be based on a sketch by Leonardo. Let's say we want to extract an outline of the birds: 

```{r birds}
im <- load.image(system.file('extdata/Leonardo_Birds.jpg',package='imager'))
plot(im)
im.g <- grayscale(im)
plot(im.g)
```

We could try thresholding:

```{r thresholding,fig.width=18}
layout(t(1:3))
threshold(im.g,"20%") %>% plot
threshold(im.g,"15%") %>% plot
threshold(im.g,"10%") %>% plot
```

The problem is that there is a luminance trend in the image (it is noticeably lighter at the bottom), so that a fixed threshold doesn't work very well. Fortunately we can remove that trend using a linear model. We're going to use R's built-in lm function. The first step is to convert the image data to a data.frame, then fit a model:

```{r remove_trend_A}
df <- as.data.frame(im.g)
head(df,5) 
m <- lm(value ~ x + y,data=df) #linear trend
summary(m)
```

We can extract fitted values to remove the luminance trend:

```{r remove_trend_B,fig.width=12}
layout(t(1:2))
im.f <- im.g-fitted(m)
plot(im.g,main="Before")
plot(im.f,main="After trend removal")
```

It's now easier to threshold the image:

```{r threshold2,fig.width=18}
layout(t(1:3))
l_ply(paste0(c(20,15,10),"%"),function(v) threshold(im.f,v) %>% plot(main=v))
```

We'll go on with a thresholding at 16%, and we can start to play with some morphological operations. "Morphological closing" will help get rid of some of the blotches: 

```{r mclosing,fig.width=12}
im.t <- threshold(im.f,"16%")
layout(t(1:2))
plot(im.t,main="Before")
mclosing_square(im.t,4) %>% plot(main="After morphological closing")
```

